Journey: A Study in Unhelpful Titles (TODO: should we get a new title?)

### 2017-04-27 paper.tex
Edits in the results section

Figure \ref{fig:mmap_latency} shows for various microbenchmarks the time taken
per mmap operation as it varies with the total amount memory allocate so far in
the system. As described above, we use rdtsc to find the number of ticks taken
for the operation and then calibrate rdtsc to conver these ticks into
microseconds. It can be seen that for continuous mmaps the time is not only the
lowest but also invariant with the total number of mmaps done so far. This is
because of two reasons. First, when a continuous mmap (with the same
permissions) happens the kernel memory allocator does not need create a new vm
area struct, instead it modifies the adjacent vm area structs to account for the
newly accommodate memory. This ensures that there is no need to perform the
lookup on the red black trees that store the vm area sturct so a horizontal line
is seen. The second important thing to note here is that for continuous mmap
operations, the time taken is always the lowest compared to other benchmarks.
This is also a result of expanding the existing vm area sturcts. Since the
number of vm area structs is very low (actually just 1 for the process in
question) the required node in the red black tree is always found in the cache.
For rest of the microbench marks, the kernel has to first search through the
red-black tree of vm area structs. If it finds a mapping adjacent to the new
memory allocation to be done, it will extend the existing mapping otherwise it
will create a new mapping and insert it into the red-black tree. The traversal
of the tree is an o(long n) operation where n is the number of nodes in the tree
which in this case are the vm area mappings. This log n shape for the lookup is
visible in the plots for all the benchmarks in figure \ref{fig:mmap_ latency}.
The frag mmap benchmark takes the most time because for fragmented mmap, the
memory is first allocated contiguously and then all allocated regions are
expanded to a higher size. This operation needs creation of a new memory area
struct. Therefore, the kernel first looks through the red-black tree but it
never finds a node which can be expanded to include the current mapping. So it
always has to create a new mapping after traversing the tree and then inserting
it in. It also has to delete the old mapping from the red-black tree and then
rebalnce the tree if necessary. All of these operations result in the worst case
latency for mmap.  The case of strided mmaps is very similar to frag because it
will also force the kerenl to always create a new mapping and insert it into the
tree after traversing the tree first. It however does not need to delete an
existing mapping so it is faster than the frag benchmark. Finally, rand will
also search through the red-black tree first but it will sometimes find an
existing mapping adjacent to the desired mapping which can be expanded obviating
the need to create a new vm area mapping and inserting it into red-black tree.
Hence, most of the times it is faster than stride.

@@

\subsubsection{Memory}
Figure \ref{fig:mmap_mem_usage} shows the growth in memory usage of the system
for the workloads described in section \ref{ss_vm_area_struct}. The
\texttt{cont} workload coincides with the \texttt{control} workload as only the
first mmap operation allocates a \texttt{vm\_area\_struct}. The subsequent
operations are subsumed by the initial \texttt{vm\_area\_struct} as the mapped
regions are adjacent, and hence the memory usage does not increase despite the
large number of mmap operations.  The \texttt{stride} workload allocates a
\texttt{vm\_area\_struct} for each operation due to the separation of
allocations in the virtual address space. As a result, the memory usage
increases significantly as more memory is consumed with each mmap operation.
This is also the case for the \texttt{rand} workload which breaks up a single
\texttt{vm\_area\_struct} for a large area in the virtual address space into
multiple smaller areas, which results in the  allocation of a new structure for
every resize operation. The \texttt{rand} workload initially creates a new
\texttt{vm\_area\_struct} struct for each mmap operation. Eventually, these
allocations form larger contiguous allocations when two or more regions are
mapped next to each other, resulting in a single \texttt{vm\_area\_struct} being
used for that range. As a result, the memory usage begins to flatten out and
even decreases slightly as multiple \texttt{vm\_area\_struct}s are collapsed to
single structures. This is made clear by Figure \ref{fig:vm_area_struct_count}
which shows the growth \texttt{vm\_area\_struct} counts for each workload.
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/mmap_mem_usage}
    \caption{$\Delta_i$ adjusted with control for benchmarks}
    \label{fig:mmap_mem_usage}
\end{figure}

>>>>

Figure \ref{fig:mmap_latency} shows for various microbenchmarks the time taken
per \texttt{mmap} operation as it varies with the total amount memory allocated
so far in the workload.

It can be seen that for continuous \texttt{mmap}s the time is not only the
lowest but is also invariant with the total number of \texttt{mmap}s done so
far. There are two reasons for this.

First, when a continuous \texttt{mmap} (with the same permissions) happens the
kernel memory allocator does not need to create a new \texttt{vm\_area\_struct},
instead it modifies the adjacent \texttt{vm\_area\_struct}s to account for the
newly allocated space. This avoids the need to use the slab allocator to create
a new struct.

Second, for continuous \texttt{mmap} operations, the time taken is always the
lowest compared to other benchmarks.  This is because the kernel caches the most
recently used \texttt{vm\_area\_struct}, avoiding a red-black tree search. In
the continuous benchmark, this cached struct is always conveniently used.

For the other workloads, a search through the red-black tree is often needed on
most accesses. This operation takes $O(log n)$ time (where $n$ is the number of
virtual memory areas). The logarithmic behavior is visible in the plots for the
benchmarks in Figure \ref{fig:mmap_latency}.

The frag benchmark takes the most time of any benchmark. When a region of the
address space is being expanded, the kernel is forced to unmap it and map a new
area somewhere else.  This operation needs creation of a new memory area struct.
The kernel also must look through the red-black tree, though it never finds a
node which can be expanded to include the new mapping. Then, an insertion on the
tree is needed, possibly including rebalancing.

The strided workload is very similar to frag because it will also force the
kerenl to always create a new mapping and insert it into the tree after
traversing the tree first. It, however, does not need to delete an existing
mapping so it is faster than the frag benchmark.

Finally, rand will also search through the red-black tree first but there is a
reasonably high probability that it will find an existing mapping adjacent to
the desired mapping which can be expanded, obviating the need to create a new
struct and insert it into red-black tree.

TODO?

@@

\subsubsection{Memory}
Figure \ref{fig:mmap_mem_usage} shows the growth in memory usage of the system
for the workloads described in section \ref{ss_vm_area_struct}. The
\texttt{cont} workload coincides with the \texttt{control} workload as only the
first \texttt{mmap} operation allocates a \texttt{vm\_area\_struct}. The subsequent
operations are subsumed by the initial \texttt{vm\_area\_struct} as the mapped
regions are adjacent, and hence the memory usage does not increase despite the
large number of \texttt{mmap} operations.  The \texttt{stride} workload allocates a
\texttt{vm\_area\_struct} for each operation due to the separation of
allocations in the virtual address space. As a result, the memory usage
increases significantly as more memory is consumed with each \texttt{mmap} operation.
This is also the case for the \texttt{rand} workload which breaks up a single
\texttt{vm\_area\_struct} for a large area in the virtual address space into
multiple smaller areas, which results in the  allocation of a new structure for
every resize operation. The \texttt{rand} workload initially creates a new
\texttt{vm\_area\_struct} struct for each \texttt{mmap} operation. Eventually, these
allocations form larger contiguous allocations when two or more regions are
mapped next to each other, resulting in a single \texttt{vm\_area\_struct} being
used for that range. As a result, the memory usage begins to flatten out and
even decreases slightly as multiple \texttt{vm\_area\_struct}s are collapsed to
single structures. This is made clear by Figure \ref{fig:vm_area_struct_count}
which shows the growth \texttt{vm\_area\_struct} counts for each workload.
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/mmap_mem_usage}
    \caption{$\Delta_i$ adjusted with control for benchmarks}
    \label{fig:mmap_mem_usage}
\end{figure}
